\section*{Acknowledgements}
\label{sec:acknowledgement}

The past years doing my PhD program at the LRDE, EPITA was rich and intense in experience. I would like to thank all of
those that have supported me through this ordeal, which is a PhD thesis.

I would like to sincerely thank Benjamin Perret and Pascale Le Gall who accepted to review my thesis. Also, I would like
to thank Hugues Talbot, Laurent Najman, Camille Kurtz and Jo\"{e}l Falcou who agreed to be part of my jury.

I would like to address my sincere thanks to Thierry Géraud. You have recruited me and trusted me from the start. Even
though the journey was bumpy and full of happenings, your advises and remarks were always on point, accurate and
helpful. In the end


and Edwin Carlinet, my thesis director and supervisor, which
have accompanied me through this ordeal that is a PhD thesis, and for always having been there. I wish I have gotten
more help, especially in the difficult moments.

I wish especially to thank Claire Lecocq, for her support and advices that allowed to overcome this ordeal.

I would like to take the time to thank Daniela, who has accompanied me all through my thesis. You helped for my
integration at the beginning and has always helped me, notably to get through all the administrative tasks at every
step of the PhD.

Special tanks to Cl\'{e}ment, our dearest Sysadmin. You were always on point, helpful and quick to make my computer work
again! Thank you for your sysadmin jokes too! They really lift up the mood.

Also, I would like to thank all the members of the Image team, Jonathan, Joseph, Guillaume, \'{E}lodie, Nicolas and
Olivier for their joyful mood and their help all along this thesis.

Of course, I would also like to thank the people of the Spot team (automata), especially Alexander, Sven, Jim, Adrien,
Philipp, and Didier. It has been (and still is) a real pleasure to work with you.

I would like to warmly thank the PhD team, always there to share a worry, to give an advice, to drink a beer, all those
essential things a PhD student needs to survive! A big thanks to Ludovic and Duy, who have set on their own path now. I
wish you the best. A big thank you to Baptiste, Caroline, Minh, Charles, Zhou, Yizi, Antoine, Anissa and Florian for
their support, especially the jokes and political debates during the breaks. Let us redo the world again around a beer
next time!

Finally, I would like to thank my beloved Familly, for their unfailing support, from the beginning to the end of this
ordeal. You have always believed in me. Thank you for that. I would have not been able to finish without you.
