\documentclass[11pt,a4paper]{book}
%\documentclass[11pt,a4paper,draft]{book}
\usepackage[inner=3cm,textwidth=16cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
% \usepackage[numbers]{natbib}
\usepackage[mode=buildmissing]{standalone}
\usepackage{tikz}
\usepackage{float}
%\usepackage{placeins}
%\usepackage{refcheck}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{minted}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{tikz-uml}
\usepackage{threeparttable}
\usepackage[]{siunitx}
\usepackage{makecell}
\usepackage{graphbox}
\usepackage{diagbox}
\usepackage[draft]{fixme}
\usepackage{makecell}
%\usepackage{bibentry}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathabx}
\usepackage{cleveref}
%\usepackage{svg}
\usepackage{lettrine}
%\usepackage{wrapfig}
%\usepackage[backend=biber,style=alphabetic,sorting=ynt]{biblatex}
%\usepackage[backend=biber,sorting=ynt,style=numeric]{biblatex}
%\usepackage[backend=biber,sorting=ynt,style=draft,backref=true]{biblatex}
\usepackage[backend=biber,
            defernumbers=true,
            sorting=ynt,
            style=numeric,
            %style=draft,
            backref=true]{biblatex}

\DeclareBibliographyCategory{cited}
\AtEveryCitekey{\addtocategory{cited}{\thefield{entrykey}}}
\addbibresource{bibliography.bib}

% This allows for an additional line-breaking pass with the amount of
% "tolerable" white space per line increased by 1em.
% Needed for overfull hbox in bibliography
% \emergencystretch=1em
\usepackage[final]{microtype}

\usepackage[section]{placeins}

\usepackage{epstopdf}
\epstopdfDeclareGraphicsRule{.tif}{png}{.png}{convert #1 \OutputFile}
\AppendGraphicsExtensions{.tif}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\eqmark}{{\bf \(\approx\)}}

\newcommand{\eg}{e.g.}
\newcommand{\ie}{i.e.}
\newcommand{\cf}{cf.}
\newcommand{\vs}{vs.}
\newcommand{\pros}{pros.}
\newcommand{\cons}{cons.}

\definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
\definecolor{lightblue}{rgb}{0.68, 0.85, 0.9}
\definecolor{lightgreen}{rgb}{0.56, 0.93, 0.56}
\definecolor{thistle}{rgb}{0.85, 0.75, 0.85}

\newcommand{\FIXME}[1]{{\color{red}#1}}
\newcommand{\PLAGIAT}[1]{{\color{magenta}#1}}

% Reset chapter number after each part
% \makeatletter
% \@addtoreset{chapter}{part}
% \makeatother

%\renewcommand{\baselinestretch}{1}

\usetikzlibrary{3d,calc,positioning}
\tikzumlset{font=\scriptsize\tt}
\setminted{fontsize=\scriptsize}

\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Generic programming in modern C++ for Image Processing}


\author{Michaël Roynard}

%\institute{EPITA Research and Development Laboratory\\
%  \email{michael.roynard@lrde.epita.fr}
%}

\maketitle

\section*{Acknowledgement}
\label{sec:acknowledgement}


\section*{Abstract}
\label{sec:abstract}
C++ is a multi-paradigm language that enables the initiated programmer to set up efficient image processing algorithms.
This language strength comes from many aspects. C++ is high-level, so this enables developing powerful abstractions and
mixing different programming styles to ease the development. At the same time, C++ is low-level and can fully take
advantage of the hardware to deliver the best performance. It is also very portable and highly compatible which allows
algorithms to be called from high-level, fast-prototyping languages such as Python or Matlab. One of the most
fundamental aspects where C++ really shines is generic programming. Generic programming makes it possible to develop and
reuse bricks of software on objects (images) of different natures (types) without performance loss. Nevertheless,
conciliating genericity, efficiency, and simplicity at the same time is not trivial. Modern C++ (post-2011) has brought
new features that made it simpler and more powerful. In this thesis, we first explore one particular C++20 aspect: the
concepts, in order to build a concrete taxonomy of image related types and algorithms. Second, we explore another
addition to C++20, ranges (and views), and we apply this design to image processing and image types in order to solve
issues such as how hard it was to customize/tweak image processing algorithms. Finally, we explore possibilities
regarding how we can offer a bridge between static (compile-time) generic C++ code and dynamic (run-time) Python code.
We offer our own hybrid solution and benchmark its performance as well as discussing what can be done in the future with
JIT technologies. Those three axes aim to address the issue of generic programming all the while remaining efficient and
easy to use.

\bigskip

\bigskip

\noindent C++ est un langage de programmation multi-paradigme qui permet au développeur initié de mettre au point des
algorithmes de traitement d'image. La force de langage se base sur plusieurs aspects. C++ est haut-niveau, cela signifie
qu'il est possible de développer des abstractions puissantes mélangeant plusieurs styles de programmation pour faciliter
le développement. En même temps, C++ est bas-niveau et peut pleinement tirer partie du matériel pour fournir un maximum
de performances. Il est aussi portable et très compatible ce qui lui permet de se brancher à d'autres langages de haut
niveau pour le prototypage rapide tel que Python ou Matlab. Un des aspects les plus fondamentaux où le C++ brille est la
programmation générique. La programmation générique rend possible le développement et la réutilisation de brique
logiciels comme des objets (images) de différentes natures (types) sans avoir de perte au niveau performance. Néanmoins,
concilier la généricité, la performance et la simplicité d'utilisation tout en même temps n'est pas trivial. Le C++
moderne (post-2011) amène de nouvelles fonctionnalités qui le rendent plus simple et plus puissant. Dans cette thèse,
nous explorons en premier un aspect particulier du C++20 : les concepts, dans le but de construire une taxonomie des
types relatifs au traitement d'image. Deuxièmement, nous explorons une autre addition au C++20, les ranges (et les
vues), et nous appliquons ce design au traitement d'image et aux types d'image dans le but résoudre les problèmes
relatifs à la difficulté de customiser les algorithmes de traitement d'image. Enfin, nous explorons les possibilités
concernant la façon dont il est possible de construire un pont entre du code C++ générique statique (compile-time) et du
code Python dynamique (run-time). Nous fournissons une solution hybride et nous mesurons ses performances. Nous
discutons aussi les pistes qui peuvent être explorées dans le futur, notamment celles qui concernent les technologies
JIT. Ces trois axes visent à solutionner la problématique concernant la programmation générique tout en restant efficace
et accessible.

\chapter*{Résumé long (FR)}
\label{sec:long_summary}
\input{chapters/0_resume_long.tex}

\tableofcontents
\label{table.of.contents}

\listoffigures
\label{list.of.figures}

\listoftables
\label{list.of.tables}

\cleardoublepage


\part{Context}
\label{part:context}

\input{chapters/1_introduction.tex}

\cleardoublepage

\input{chapters/2_genericity.tex}

\cleardoublepage


\part{Contribution}
\label{part:contribution}

\input{chapters/3_Images_and_algorithms_taxonomy.tex}

\cleardoublepage

\input{chapters/4_images_views.tex}

\cleardoublepage

\input{chapters/5_static-dynamic_bridge.tex}


\part{Continuation}
\label{part:continuation}

\cleardoublepage

\input{chapters/6_conclusion_and_perspectives.tex}

\cleardoublepage


\part{Appendices}
\label{part:annexes}

\appendix

% This allows for an additional line-breaking pass with the amount of
% "tolerable" white space per line increased by 1em.
% Needed for overfull hbox in bibliography
\emergencystretch=1em

\chapter{Bibliography}
\label{chap:bibliography}

\nocite{*}

\printbibliography[category=cited,heading=bibintoc,title={References}]
\printbibliography[notcategory=cited,heading=bibintoc,title={Further reading}]

\cleardoublepage

\input{chapters/7_appendix_concepts_and_archetypes.tex}

\input{chapters/8_appendix_benchmark_static-dynamic_bridge.tex}


\end{document}
